Convex Surrogates via Polyhedral Losses

### 2019-04-25 neurips-19/cvx-fin.tex
typo fixes, notes

\documentclass[12pt]{article}
\usepackage[preprint]{neurips_2019}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage[implicit=false, bookmarks=false]{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\usepackage{mathtools, amsmath, amssymb, amsthm, graphicx, verbatim}
%\usepackage[thmmarks, thref, amsthm]{ntheorem}
\usepackage{color}
\usepackage{wrapfig}
\usepackage{subcaption}
\usepackage[colorinlistoftodos,textsize=tiny]{todonotes} % need xargs for below
%\usepackage{accents}
\usepackage{bbm}
\usepackage{xspace}

@@

\title{Convex Surrogates via Polyhedral Losses}
\author{%
 Jessica Finocchiaro \\
 \texttt{jefi8453@colorado.edu}\\
 CU Boulder
 \And
 Rafael Frongillo\\
 \texttt{raf@colorado.edu}\\
 CU Boulder
 \And
 Bo\\
 \texttt{bwag@colorado.edu}\\
 MSFT
}

@@

\begin{abstract}
  We formalize and investigate the natural approach of designing convex surrogate loss functions via embeddings, for problems such as classification or ranking which involve a finite number of possible predictions.
  Specifically, we consider surrogate losses given by embedding each prediction (label, ranking, etc) as a point in $\reals^d$, and convexifying the loss.
  We prove that this approach is equivalent to working with polyhedral (piecewise linear convex) losses, in several strong senses.
  We then investigate the dimensionality of the surrogate hypothesis space required for such an embedding, giving a complete characterization of the $1$-d case, and several necessary conditions in higher dimensions.
  We illustrate our results with new lower bounds for abstain loss.
\end{abstract}

@@

\begin{definition}[Finite property, non-redundant]
  A property $\Gamma:\simplex\toto\R$ is \emph{redundant} if for some $r,r'\in\R$ we have $\Gamma_r \subseteq \Gamma_{r'}$, and \emph{non-redundant} otherwise.
  $\Gamma$ is \emph{finite} if it is non-redundant and $\R$ is \jessiet{delete: a} a finite.
  We typically write finite properties in lower case, as $\gamma$.
\end{definition}

Thus, $\Gamma(p)$ is the set of reports which should be optimal for a given distribution $p$, and $\Gamma_r$ is the set of distributions for which the report $r$ should be optimal.
For example, the \emph{mode} is the property $\mode(p) = \argmax_{y\in\Y} p_y$, which is finite.
The mode captures the set of optimal reports for 0-1 loss: for each distribution over the labels, one should report the most likely label.
In this case we say 0-1 loss \emph{elicits} the mode, as we formalize below.
This terminology comes from the information elicitation \jessiet{information vs property... do we care?} literature~\citep{savage1971elicitation,osband1985information-eliciting,lambert2008eliciting}, in which a report $r$ is elicited from some forecaster by scoring her with a loss on the observed outcome $y$.

@@

Often one is given a property $\gamma$ (typically via a discrete loss $\ell$) and \jessiet{delete: one} one seeks a surrogate loss $L$ with certain attributes, such as convexity, smoothness, etc.
In doing so, we may be interested in the minimum number of dimensions needed for such an indirect elicitation scheme; this is the notion of \emph{elicitation complexity}~\citep{lambert2008eliciting,fissler2016higher,frongillo2015elicitation}.
The two notions of elicitation complexity we use most are as follows: let $\eliccvx(\gamma)$ be the minimum dimension $d$ such that $\gamma$ can be indirectly elicited by a convex loss function $L: \reals^d \to \reals^\Y_+$, and $\elicpoly(\gamma)$ the minimum such $d$ where $L$ is polyhedral.
We will also discuss $\eliccts$, which is the complexity with respect to continuous, real-valued, non-locally-constant properties (\S~\ref{sec:1d}), and $\elicembed$, which is the minimum dimension for $\gamma$ to be embedded (Definition~\ref{def:embedding-dim}).

@@

\begin{definition}
  A property $\Gamma : \simplex \toto \reals^d$ \emph{embeds} a property $\gamma : \simplex \toto \R$ if there exists some injective embdedding $\varphi:\R\to\reals^d$ such that for all $p\in\simplex,r\in\R$ we have $r \in \gamma(p) \iff \varphi(r) \in \Gamma(p)$.
  Similarly, we say a loss $L:\reals^d\to\reals^\Y$ embeds $\gamma$ if $\prop{L}$ embeds $\gamma$.
\end{definition}

@@

With a solid understanding of the structure of embeddings and embeddable properties, we now turn to the relationship between embeddings and polyhedral losses.
We give a tight connection: every finite elicitable property is embedded by a polyhedral loss, and every polyhedral loss embeds a finite elicitable property.
We state this equivalence in two parts, the first focusing on the property, and the second on the loss.
\jessiet{I'm guessing the statement changed after this was written, but three parts?}

@@

We now state a converse of Theorem~\ref{thm:finite-elicit-embed}: every polyhedral loss embeds a finite elicitable property.
One way to look at this result in light of Proposition~\ref{prop:embed-trim} is that polyhedral losses ``directly'' elicit finite properties by embedding them into a space where the original loss is convex.
Moreover, we show that indirect elicitation via polyhedral losses are heavily structured: if a polyhedral loss elicits a property $\gamma'$ with a link to $\gamma$, then both $\gamma'$ and $\gamma$ are finite and one ``refines'' the other in the sense below.
\jessiet{Is $\gamma'$ actually finite by our definition, or is it ``finite-ish'', i.e. finite if we ignore a null set of distributions?}
In other words, the cells (level sets) of $\gamma'$ must be a finite subdivision of the cells of $\gamma$.
This fact will become useful when we discuss the relationship between embedding dimension and elicitation complexity.\raf{LATER: see [WHERE WE DISCUSS]}

@@

%\end{proof}
\newpage
\section{To Sort}
\jessiet{Lol we should probably change the name of this}
Several definitions from \citet{aurenhammer1987power}.
\begin{definition}\label{def:cell-complex}
  A \emph{cell complex} in $\reals^d$ is a set $C$ of faces (of dimension $0,\ldots,d$) which (i) union to $\reals^d$, (ii) have pairwise disjoint relative interiors, and (iii) any nonempty intersection of faces $F,F'$ in $C$ is a face of $F$ and $F'$ and an element of $C$.
\end{definition}

@@

\begin{theorem}\label{thm:general-duality-embedding}
  Let $\gamma:\simplex\toto\R$ be a non-redundant elicitable property, elicited by a loss whose Bayes risk is a closed function.
  Then $\gamma$ is $(n-1)$-embeddable, where $n=|\Y|$.
  \raft{Low priority: what conditions on $L$ or $\gamma$ ensure that the Bayes risk is closed?}
  \jessiet{\url{https://en.wikipedia.org/wiki/Closed_convex_function} so the simplex $\simplex$  being closed should suggest the Bayes Risk and $G$ are closed.}
\end{theorem}
\begin{proof}
  $G:\simplex\to\reals$ be the negative Bayes risk of the loss $L_\gamma$ eliciting $\gamma$, given by $G(p) = -\min_{r\in\R} \inprod{p}{L_\gamma(r)}$.
  Let $C:\reals^\Y\to\reals$ be the convex conjugate of $G$, given by $C(u) = \sup_{p\in\simplex} \inprod{p}{u} - G(p)$, which is finite on all of $\reals^\Y$ by~\citet[Corollary 13.3.1]{rockafellar1997convex} as $\dom\, G = \simplex$ is bounded.

>>>>

\documentclass[12pt]{article}
\PassOptionsToPackage{numbers, compress}{natbib}
\usepackage[preprint]{neurips_2019}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks  %[implicit=false, bookmarks=false]
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\usepackage{mathtools, amsmath, amssymb, amsthm, graphicx, verbatim}
%\usepackage[thmmarks, thref, amsthm]{ntheorem}
\usepackage{color}
\definecolor{darkblue}{rgb}{0.0,0.0,0.2}
\hypersetup{colorlinks,breaklinks,
            linkcolor=darkblue,urlcolor=darkblue,
            anchorcolor=darkblue,citecolor=darkblue}
\usepackage{wrapfig}
\usepackage{subcaption}
\usepackage[colorinlistoftodos,textsize=tiny]{todonotes} % need xargs for below
%\usepackage{accents}
\usepackage{bbm}
\usepackage{xspace}

@@

\title{Convex Surrogates via Polyhedral Losses}
\author{%
 Jessica Finocchiaro\raf{Jessie?} \\
 \texttt{jessica.finocchiaro@colorado.edu}\\
 CU Boulder
 \And
 Rafael Frongillo\\
 \texttt{raf@colorado.edu}\\
 CU Boulder
 \And
 Bo\\
 \texttt{bwag@colorado.edu}\\
 MSFT
}

@@

\begin{abstract}
  We formalize and investigate the natural approach of designing convex surrogate loss functions via embeddings, for problems such as classification or ranking which involve a finite number of possible predictions.
  Specifically, we consider surrogate losses given by embedding each prediction (label, ranking, etc) as a point in $\reals^d$, and convexifying the loss.
  We prove that this approach is equivalent to working with polyhedral (piecewise linear convex) losses, in several strong senses.
  We then investigate the dimensionality of the surrogate hypothesis space required for such an embedding, giving a complete characterization of the 1-dimensional case.
  We illustrate our results with modest lower bounds for abstain loss, and conclude with future directions.
\end{abstract}

@@

\begin{definition}[Finite property, non-redundant]
  A property $\Gamma:\simplex\toto\R$ is \emph{redundant} if for some $r,r'\in\R$ we have $\Gamma_r \subseteq \Gamma_{r'}$, and \emph{non-redundant} otherwise.
  $\Gamma$ is \emph{finite} if it is non-redundant and $\R$ is a finite set.
  We typically write finite properties in lower case, as $\gamma$.
\end{definition}

Thus, $\Gamma(p)$ is the set of reports which should be optimal for a given distribution $p$, and $\Gamma_r$ is the set of distributions for which the report $r$ should be optimal.
For example, the \emph{mode} is the finite property $\mode(p) = \argmax_{y\in\Y} p_y$, and captures the set of optimal reports for 0-1 loss: for each distribution over the labels, one should report the most likely label.
In this case we say 0-1 loss \emph{elicits} the mode, as we formalize below.
This terminology comes from the information elicitation \jessiet{information vs property... do we care?}\raft{we could also say economics literature; not ``property'' though since that doesn't offer an explanation of the word ``elicits''} literature~\citep{savage1971elicitation,osband1985information-eliciting,lambert2008eliciting}, in which a report $r$ is elicited from some forecaster by scoring her with a loss on the observed outcome $y$.

@@

Often one is given a property $\gamma$ (typically via a discrete loss $\ell$) and seeks a surrogate loss $L$ with certain attributes, such as convexity, smoothness, etc.
In doing so, we may be interested in the minimum number of dimensions needed for such an indirect elicitation scheme; this is the notion of \emph{elicitation complexity}~\citep{lambert2008eliciting,fissler2016higher,frongillo2015elicitation}.
The two notions of elicitation complexity we use most are as follows: let $\eliccvx(\gamma)$ be the minimum dimension $d$ such that $\gamma$ can be indirectly elicited by a convex loss function $L: \reals^d \to \reals^\Y_+$, and $\elicpoly(\gamma)$ the minimum such $d$ where $L$ is polyhedral.
We will also discuss $\eliccts$, which is the complexity with respect to continuous, real-valued, non-locally-constant properties (\S~\ref{sec:1d}), and $\elicembed$, which is the minimum dimension for $\gamma$ to be embedded (Definition~\ref{def:embedding-dim}).

@@

\begin{definition}
  A property $\Gamma : \simplex \toto \reals^d$ \emph{embeds} a property $\gamma : \simplex \toto \R$ if there exists some injective embdedding $\varphi:\R\to\reals^d$ such that for all $p\in\simplex,r\in\R$ we have $r \in \gamma(p) \iff \varphi(r) \in \Gamma(p)$.
  Similarly, we say a loss $L:\reals^d\to\reals^\Y$ embeds $\gamma$ if $\prop{L}$ embeds $\gamma$.

  Finally, a loss $L:\reals^d\to\reals^\Y$ \emph{embeds} a loss $\ell:\R\to\reals^\Y$ if $\prop{L}$ embeds $\prop{\ell}$ via $\varphi$ and for all $r\in\R$ we have $L(\varphi(r)) = \ell(r)$.
\end{definition}

\raf{We could potentially shortcut much of this.
  \\[4pt]
  A loss $L:\reals^d\to\reals^\Y$ \emph{embeds} a loss $\ell:\R\to\reals^\Y$ if there exists some injective embdedding $\varphi:\R\to\reals^d$ such that
  (1) for all $r\in\R$ we have $L(\varphi(r)) = \ell(r)$, and (2) for all $p\in\simplex,r\in\R$ we have \[r \in \argmin_{r'\in\R} \inprod{p}{\ell(r')} \iff \varphi(r) \in \argmin_{u\in\reals^d} \inprod{p}{L(u)}\]
  \\[4pt]
Maybe even simplify further?  This could help us introduce the concept before we get into properties and elicitation at all.}

@@

With a solid understanding of the structure of embeddings and embeddable properties, we now turn to the relationship between embeddings and polyhedral losses.
We give a tight connection: every finite elicitable property is embedded by a polyhedral loss, and every polyhedral loss embeds a finite elicitable property.
We state this equivalence in two theorems, the first focusing on the property, and the second on the loss.
\jessiet{I'm guessing the statement changed after this was written, but three parts?}
\raft{changed to ``theorems'' to be clearer}

@@

We now state a converse of Theorem~\ref{thm:finite-elicit-embed}: every polyhedral loss embeds a finite elicitable property.
One way to look at this result in light of Proposition~\ref{prop:embed-trim} is that polyhedral losses ``directly'' elicit finite properties by embedding them into a space where the original loss is convex.
Moreover, we show that indirect elicitation via polyhedral losses are heavily structured: if a polyhedral loss elicits a property $\gamma'$ with a link to $\gamma$, then both $\gamma'$ and $\gamma$ are finite and one ``refines'' the other in the sense below.
\jessiet{Is $\gamma'$ actually finite by our definition, or is it ``finite-ish'', i.e. finite if we ignore a null set of distributions?}
\raft{Good call, $\gamma'$ will not be finite... we'll have to revisit this whole section anyway though.}
In other words, the cells (level sets) of $\gamma'$ must be a finite subdivision of the cells of $\gamma$.
This fact will become useful when we discuss the relationship between embedding dimension and elicitation complexity.\raf{LATER: see [WHERE WE DISCUSS]}

@@

%\end{proof}
\newpage
\section{To Sort}
\jessiet{Lol we should probably change the name of this}
\raft{No kidding!  Was this really in the COLT submission??  Hilarious.}
Several definitions from \citet{aurenhammer1987power}.
\begin{definition}\label{def:cell-complex}
  A \emph{cell complex} in $\reals^d$ is a set $C$ of faces (of dimension $0,\ldots,d$) which (i) union to $\reals^d$, (ii) have pairwise disjoint relative interiors, and (iii) any nonempty intersection of faces $F,F'$ in $C$ is a face of $F$ and $F'$ and an element of $C$.
\end{definition}

@@

\begin{theorem}\label{thm:general-duality-embedding}
  Let $\gamma:\simplex\toto\R$ be a non-redundant elicitable property, elicited by a loss whose Bayes risk is a closed function.
  Then $\gamma$ is $(n-1)$-embeddable, where $n=|\Y|$.
  \raft{Low priority: what conditions on $L$ or $\gamma$ ensure that the Bayes risk is closed?}
  \jessiet{\url{https://en.wikipedia.org/wiki/Closed_convex_function} so the simplex $\simplex$  being closed should suggest the Bayes Risk and $G$ are closed.}\raft{It's more complicated than that though -- e.g. take a continuous convex function on the simplex and and then increase the values of the function arbitrarily on the poles of the simplex.  Still convex, but not closed (or continuous)}
\end{theorem}
\begin{proof}
  $G:\simplex\to\reals$ be the negative Bayes risk of the loss $L_\gamma$ eliciting $\gamma$, given by $G(p) = -\min_{r\in\R} \inprod{p}{L_\gamma(r)}$.
  Let $C:\reals^\Y\to\reals$ be the convex conjugate of $G$, given by $C(u) = \sup_{p\in\simplex} \inprod{p}{u} - G(p)$, which is finite on all of $\reals^\Y$ by~\citet[Corollary 13.3.1]{rockafellar1997convex} as $\dom\, G = \simplex$ is bounded.
